Texture mapping 3D planar models of indoor environments with noisy camera poses

### 2012-12-30 spie/paper.tex
few more holes to fill

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems often suffer from inaccuracies, resulting in visible
  discontinuities when successive images are projected adjacently onto
  a surface for texturing. Existing methods to stitch images together
  are often computationally expensive and work independently of pose
  estimates and geometry information. We propose a method to refine
  camera poses using both existing estimates and geometry information,
  followed by two different methods to composite images together,
  based on the uniformity of available images. The effectiveness of
  our methods is demonstrated on a number of different indoor
  environments.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time-consuming, and as a
result, automated 3D site modeling has garnered much interest in
recent years. CITATION

The first step in automated 3D modeling is the physical scanning of an
environment's geometry. An indoor modeling system must recover its own
poses within an environment while simultaneously reconstructing the 3D
structure of the environment itself. CITATION. This is known as the
simultaneous localization and mapping (SLAM) problem, and is generally
solved by taking readings from laser range scanners, cameras, and
inertial measurement units (IMUs) at multiple locations within the
environment.

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in much higher localization error and
uncertainty overall. CITATION. As a result, common methods for texture
mapping generally produce poor results, leading to the development of
the approaches contained in this paper. Besides localization
information and images, environment geometry is required as well for
texture mapping. In this paper we work mainly with models obtained by
fitting planar surfaces to point clouds generated by the backpack
system. CITATION. We have also developed an appropriate method to work
with near-planar triangular meshes, as described in APPENDIX
SOMETHING.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is easily generalizable to any other
system, provided the required inputs of planar geometry, images with
rough extrinsic pose estimates, and either known or estimated
intrinsic camera calibration matrices. An important benefit of our
method is its relatively low complexity and modularity of each step,
allowing for quick tuning of parameters and easy accomodation of new
input data given updates in localization or model-generation
algorithms.

The remainder of the paper is organized as follows. Section X explains
how input images are projected onto our geometry and simple approaches
towards texturing. Section X covers more sophisticated existing
approaches to image matching and stitching, and their performance on
our datasets. Section X contains our approach towards image alignment,
followed by Section X, which describes compositing approaches. Section
X contains further results and conclusions, as well as areas for
future work.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simple Texture Mapping}
\label{sec:simpleTextureMapping}
\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera matrics
    $P_{1..m}$. }
  \label{fig:projection}
\end{figure}

@@

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
in our case 5 pixels across, and choosing an image to texture each
tile.

@@

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.jpg}
  \caption{We minimize camera angle $\alpha$ and distance $d$ by
    maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
    \cdot \vec{n}$}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}[h!]
  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_naive.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_naive_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_cache_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_cache_shift_blend.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.97in]{wall2_shortest_shift_blend.pdf}}
  \caption{(a) Simple mapping. (b) Simple mapping after image
    alignment. (c) Mapping with caching after image alignment. (d)
    Shortest path texturing after image alignment). (e) same as (c)
    with blending. (f) same as (d) with blending.}
  \label{fig:compareAll}
\end{figure}

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{Graph_crop.pdf}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}
  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm from \cite{chen2010indoor} and (b) image
    mosaicing.}
  \label{fig:mosaic3D}
\end{figure}

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom. CITATION. Applying the approach in the LIU2010INDOOR paper
however, results again in error accumulation and drift as shown in
Figure \ref{fig:mosaic3D}(a). Furthermore, such nonlinear optimization
approaches IS IT? BETTER CHECK have high complexity and runtime, and the
entire solution must be recomputed given any changes in input data.

\section{2D Image Alignment with Geometry Information}
\label{sec:2dAlignment}
In this section, we describe our method of efficient and robust image
alignment. Our approach consists of four parts. First, all images are
projected onto the surface and rotated such that detected lines within
them are parallel to lines comprising the surface's boundary and
intersection with other surfaces. Next, these projections are shifted
such that interior lines near surface intersections are matched
up. Following this, occlusion checks are performed to remove invalid
parts of each image for the target surface. These first three steps
are all image-independent and are performed in parallel. Finally, we
detect SIFT feature matches between pairs of images and solve a linear
least squares problem in 2D to maximize matches.

All of our calculations and alignments are performed in 2D, partly for
efficiency, and partly because the nature of our input data is such
that the majority of localization error is 2D, in the same plane as
the surface being projected onto. Recall that our data acquisition
system is backpack-mounted, with cameras facing to the sides of the
operator. The operator attempts to walk parallel to walls, and the
localization and model-generation algorithms that provide our input
can conform the operator's path to be straight and parallel to
detected surfaces. Since the operator walks as upright as possible,
errors in roll and yaw, as well as the operator's distance from
parallel surfaces are minimal. Thus, our highest errors stem from
uncertainty in the operator's pitch, equivalent to rotation around the
camera axes, as well as location along surfaces. These equate to 2D
rotation and translation in a surface's plane.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect lines
in the projections using Hough transforms. Experience and intuition
show that walls in indoor environments contain linear features that
are either horizontal or vertical, often corresponding to doors,
windows, posters, etc. Thus, when texturing walls, we rotate images in
2D such that dominant lines are made to be horizontal or
vertical. Though not relevant in our datasets, this can be
extrapolated such that lines are oriented with a wall's boundaries,
for example in areas with a slanted roof.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{geometryFeature.jpg}
  \includegraphics[height=1.5in]{imageCorner.jpg}
  \includegraphics[height=1.5in]{geometryImageAlign.jpg}
  \caption{(a) The red lines indidcate plane intersections that can be used as geometry reference points. (b) A corresponding corner in an image, detected as a line. (c) The image rotated and shifted to align with geometry.}
  \label{fig:geometryAlignment}
\end{figure}

@@

To perform the rotation and alignment, a robust method would be to
detect all line matches and solve for an optimal homography. In nearly
all cases however, we find that a single 2D rotation is sufficient to
properly align lines in an image. We also rarely have more than one
linear geometry feature to match to. As a result, we do WHATEVER WE DO, and then indicate which images are aligned to geometry, which is useful in Section \ref{sec:robustSIFTFeatureMatching}

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, nothing occurs. Where there is
a mixture of both, the rectangular region is subdivided into four, and
the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion, we
remove all erroneous texture belonging to other surfaces, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lai1999robust, lowe1999object, mikolajczyk2005performance,
  szeliski2006image}.  In our implementation, we use SiftGPU, by
PERSON CITATION, for both feature detection as well as pairwise
matching. These matches determine $d^x$ and $d^y$ distances between
each pair of features for two image projections, though these
distances may not always be the same for different features. Since
indoor environments often contain repetitive features such as floor
tiles or doors, we need to ensure that SIFT-based distances are
reliable. First, we only perform alignment on pairs of images that
overlap given the original noisy poses. Secondly, we discard feature
matches that correspond to an image distance greater than 40 pixels
from what the noisy poses estimate. In order to utilize the remaining
feature matches robustly, the RANSAC framework
\cite{fischler1981random} is used to estimate the optimal $d^x_{i,j}$
and $d^y_{i,j}$ distances between two images $i$ and $j$. The RANSAC
framework requires a fitting function and a distance function. For
this application, the fitting function simply finds the average
distance between matches in a pair of images. The distance function
for a pair of points is chosen to be the difference between those
points' SIFT match distance and the average distance computed by the
fitting function. We use a 10 pixel outlier threshold, so that SIFT
matches are labeled as outliers if their horizontal or vertical
distances are not within 10 pixels of the average distance computed by
the fitting function.

We now use the calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. Recall that there are a total of $M^{2}$
possible pairs of images, though we only generate distances between
images that overlap at SIFT feature points. Given these distances and
the original image location estimates, we can solve a least squares
problem ($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} -
\vec{\gamma}||_2^2 $) to estimate the location of the images on the
plane. The $M$-dimensional vector $\vec{\beta}$ represents the unknown
$x$ location of each image on the plane for $1 \dots M$. The optimal
$x$ and $y$ locations are obtained in the same way, so we only
consider the $x$ locations here:

@@

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates a SIFT-based distance between images 1 and 2,
images 1 and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]
If only relative distances between images are included, the absolute
location of the images can not be calculated, and the matrix is rank
deficient. From Section \ref{geometryBasedAlignemtn}, a number of
images are anchored to geometry points, and thus their locations can
be used to anchor the rest in place. In case no anchor images exist,
we simply arbitrarily pick an image. In the above matrix, the first
image is set to be such an anchor, simply by placing a $1$ in its
column

The $N$-dimensional observation vector $\vec{\gamma}$ is constructed
using the SIFT-based distances generated in the RANSAC matching
stage. Elements in the observation vector corresponding to anchor
images are simply their locations as determined by the original noisy
localization. Thus $\vec{\gamma}$ can be written as:

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. This solution however does not make use of our noisy camera
poses, and will fail when no SIFT matches are found between one
segment of the plane and another. To account for this, we add rows to
the $A$ matrix and observations to the $\vec{\gamma}$ vector
corresponding to the original noisy distances. We then solve a
weighted least squares problem where the SIFT distances and anchor
values are given a high weight e.g. 1, while the noisy distances are
given a smaller weight e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:simpleTextureMapping} results in X, which has far
fewer seams, but can still be improved.

\section{Image Compositing}
\label{sec:imageCompositing}
In this section, we revisit the tile-based texturing approach from
Section \ref{sec:simpleTextureMapping}, with an added caching
mechanism to reduce image boundaries. This method works well given all
manner of camera poses and surfaces, but for cases where we have large
sections of usable texture from images, we propose an alternate method
that further reduces image boundaries.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
From Section \ref{sec:simpleTextureMapping}, we saw that
discontinuities occur where adjacent tiles are textured by different
images. Though image alignment removes many such discontinuities,
there are still cases where seams are visible due to imprecise
matching or other factors such as lighting differences. To reduce the
cases where this happens, it makes sense to take into account image
selections made by neighboring tiles while texture mapping a given
tile. By using the same image across tile boundaries, we can eliminate
a discontinuity altogether. If this is not possible because a tile is
not visible in images used by neighboring tiles, using similar images
across tile boundaries also leads to less noticeable discontinuities.

@@

If no satisfactory image is found in the first subset, we check the
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but are taken at a
similar location and time, suggesting that their localization and
projection are quite similar, and thus likely matched more
accurately. Again, if no viable image is found according to the same
criteria, we search the entire set of candidate images, selecting
based on the same scoring function from Figure
\ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure X, where seams
are now greatly reduced as compared to Figure X2.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Images for vertical planes are tilted, but their
    effective camera axis is more or less normal to the plane. (b) The
    effective camera axis for ceilings is at an angle with respect to
    the plane normal. (c) Wall images are cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

>>>>

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}

  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems often suffer from inaccuracies, resulting in visible
  discontinuities when successive images are projected adjacently onto
  a surface for texturing. Existing methods to stitch images together
  are often computationally expensive and work independently of pose
  estimates and geometry information. We propose a method to refine
  camera poses using both existing estimates and geometry information,
  followed by two different methods to composite images together,
  based on the uniformity of available images. The effectiveness of
  our methods is demonstrated on a number of different indoor
  environments.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section

Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time-consuming, and as a
result, automated 3D site modeling has garnered much interest in
recent years.

The first step in automated 3D modeling is the physical scanning of an
environment's geometry. An indoor modeling system must recover its own
poses within an environment while simultaneously reconstructing the 3D
structure of the environment itself. \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure}. This is known as the
simultaneous localization and mapping (SLAM) problem, and is generally
solved by taking readings from laser range scanners, cameras, and
inertial measurement units (IMUs) at multiple locations within the
environment.

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in much higher localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results, leading to the development of the
approaches contained in this paper. Besides localization information
and images, environment geometry is required as well for texture
mapping. In this paper we work mainly with models obtained by fitting
planar surfaces to point clouds generated by the backpack system
\cite{sanchez2012point}. We have also developed an appropriate method
to work with near-planar triangular meshes, as described in Appendix
A.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is applicable to other systems,
provided the required inputs of planar geometry, images with rough
extrinsic pose estimates, and either known or estimated intrinsic
camera calibration matrices. An important benefit of our method is its
relatively low complexity and runtime, and independence of many
steps, allowing for quick tuning of parameters and easy
accomodation of new input data given updates in localization or
model-generation algorithms.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how input images are projected
onto our geometry and simple approaches towards texturing. Section
\ref{sec:existingApproaches} covers more sophisticated existing
approaches to image matching and stitching, and their performance on
our datasets. Section \ref{sec:imageAlignment} contains our approach
towards image alignment, followed by Section
\ref{sec:imageCompositing}, which describes compositing
approaches. Section \ref{sec:resultsAndConclusions} contains further
results and conclusions, as well as areas of ongoing and future work.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simple Texture Mapping}
\label{sec:simpleTextureMapping}

@@

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera matrics
    $P_{1..m}$. }
  \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.jpg}
  \caption{We minimize camera angle $\alpha$ and distance $d$ by
    maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
    \cdot \vec{n}$}
  \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
in our case 5 pixels across, and choosing an image to texture each
tile.

@@

\begin{figure}
  \centering 
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_naive.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_naive_shift.pdf}}

  \centering
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_cache_shift.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_shortest_shift.pdf}}

  \centering 
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_shortest_shift_blend.pdf}}
  \caption{(a) Tile-based texturing. (b) Tile-based texturing after
    image alignment. (c) Tile-based texturing after image alignment
    with caching. (d) Shortest path texturing after image
    alignment). (e) Same as (c) with blending. (f) Same as (d) with
    blending.}
  \label{fig:compareAll}
\end{figure}

@@

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{Graph_crop.pdf}}
  \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}

  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm from \cite{chen2010indoor} and (b) image
    mosaicing.}
  \label{fig:mosaic3D}
\end{figure}

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom \cite{liu2010indoor}. The approach in the cited paper works
well locally, given high feature density, but over larger areas
results again in error accumulation and drift, especially in
feature-sparse areas, as shown in Figure
\ref{fig:mosaic3D}(a). Furthermore, it adjusts image locations using
an iterative process, which results in high runtime.

\section{2D Image Alignment with Geometry Information}
\label{sec:2dAlignment}
In this section, we describe our method of efficient and robust image
alignment. Our approach consists of three main parts. First, all
images are projected onto the surface and lines within these
projections are detected. These lines are then matched up with
geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a linear least squares problem in 2D to maximize
matches.

All of our following calculations and alignments are performed in 2D,
partly for efficiency, and partly because the nature of our input data
is such that localization error chiefly occurs in two dimensions, in
the same plane as the surface being projected onto. Recall that our
data acquisition system is backpack-mounted, with cameras facing to
the sides of the operator. The operator makes efforts to walk parallel
to walls, and the localization and model-generation algorithms that
provide our input often conform the operator's path to be straight and
parallel to detected surfaces \cite{kua2012loopclosure,
  sanchez2012point}. Since the operator walks in as upright a position
as possible, errors in roll, yaw, and the operator's distance from
parallel surfaces are minimal. Thus, our highest errors stem from
uncertainty in the operator's pitch, equivalent to rotation around the
camera axes, as well as location along surfaces. These equate to 2D
rotation and translation in a surface's plane.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect lines
in the projections using Hough transforms. Experience and intuition
show that walls in indoor environments contain linear features that
are either horizontal or vertical, often corresponding to doors,
windows, posters, etc. Thus, when texturing walls, we rotate images in
2D such that dominant lines are made to be horizontal or
vertical. This can be further extrapolated by orienting lines with a
wall's boundaries, for example in areas with a slanted roof.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{geometryFeature.jpg}
  \includegraphics[height=1.5in]{imageCorner.jpg}
  \includegraphics[height=1.5in]{geometryImageAlign.jpg}
  \caption{(a) The red lines indidcate geometry-based lines,
    consisting of boundaries and intersections. (b) A matching
    image-based line, corresponding to a corner. (c) The image rotated
    and shifted such that the matched lines align align with
    geometry.}
  \label{fig:geometryAlignment}
\end{figure}

@@

To perform the rotation and alignment, one method would be to detect
all line matches and solve for an optimal homography. In nearly all
cases however, we find that a single 2D rotation is sufficient to
properly align lines in an image. We also rarely have more than one
linear geometry feature to match to in each orthogonal
direction. Thus, for a rotation angle, we simply take all angles
between image-based and geometry-based lines under 5 degrees, and
average them. After the image is rotated, we find the horizontal and
vertical image-based and geometry-based line segments with minimum
orthogonal distance between them. For both distances, if it is under 5
mm, we apply the corresponding shift, and indicate that the image was
aligned to geometry in one or both dimensions, which reduces further
shifting in Section \ref{sec:robustSIFTFeatureMatching}.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, nothing occurs. Where there is
a mixture of both, the rectangular region is subdivided into four, and
the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
pairs of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, the RANSAC framework
\cite{fischler1981random} is used to estimate the optimal $d^x_{i,j}$
and $d^y_{i,j}$ distances between two images $i$ and $j$. The RANSAC
framework requires a fitting function and a distance function. For
this application, the fitting function finds the average distance
between matches in a pair of images, and the distance function for a
pair of points is chosen to be the difference between those points'
SIFT match distance and the average distance computed by the fitting
function. We use a 10 pixel threshold, so that SIFT matches are
labeled as outliers if their horizontal or vertical distances are not
within 10 pixels of the average distance computed by the fitting
function.

We now use the calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. There are a total of $M^{2}$ possible pairs of
images, though we only generate distances between images with SIFT
matches. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

@@

The $N \times M$ matrix $A$ is constructed with one row for each pair
of images with measured distances produced by the SIFT matching
stage. A row in the matrix has a $-1$ and $1$ in the columns
corresponding to the two images in the pair. For example, the matrix
below indicates a SIFT-based distance between images 1 and 2, images 1
and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0\\
\end{pmatrix}
\]
If only relative distances between images are included, the absolute
location of the images can not be calculated, and the matrix is rank
deficient. From Section \ref{geometryBasedAlignemtn}, a number of
images are anchored to geometry points, in one or both dimensions, and
thus their locations can be used to fix the rest in place. In case no
anchor images exist, we simply arbitrarily pick an image, as we have
no other reference points to work with. In the above matrix, the first
image is set to be such an anchor, simply by placing a $1$ in its
column.

The $N$-dimensional observation vector $\vec{\gamma}$ is constructed
using the SIFT-based distances generated in the RANSAC matching
stage. Elements in the observation vector corresponding to anchor
images are simply their locations as determined by the original noisy
localization. Thus a $\vec{\gamma}$ corresponding to the above matrix
can be written as:

@@

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. This solution however does not make use of our noisy camera
poses, and will fail when no SIFT matches are found between one
segment of the plane and another. To account for this, we add rows to
the $A$ matrix and observations to the $\vec{\gamma}$ vector
corresponding to the original noisy distances. We then add weighting
to our least squares problem where the SIFT distances and anchor
values are given a high weight e.g. 1, while the noisy distances are
given a smaller weight e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that section results in Figure \ref{fig:compareAll}(b), which has far
fewer seams, though some are still visible.

\section{Image Compositing}
\label{sec:imageCompositing}
In this section, we revisit the tile-based texturing approach from
Section \ref{sec:simpleTextureMapping}, with an added caching
mechanism to reduce image boundaries. This method works well given all
manner of camera poses and surfaces, but for optimal cases where we
have large sections of usable texture from images, we propose a
superior method that further reduces image boundaries.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
From Section \ref{sec:tileBasedMapping}, we saw that discontinuities
occur where adjacent tiles are textured by different images. Though
Section \ref{sec:2dAlignment}'s image alignment removes many such
discontinuities, there are still cases where seams are visible due to
imprecise matching or other factors such as model-based errors. To
reduce the cases where this happens, it makes sense to take into
account image selections made by neighboring tiles while texture
mapping a given tile. By using the same image across tile boundaries,
we can eliminate a discontinuity altogether. If this is not possible
because a tile is not visible in images used by neighboring tiles,
using similar images across tile boundaries also leads to less
noticeable discontinuities.

@@

If no satisfactory image is found in the first subset, we check a
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but are taken at a
similar location and time, suggesting that their localization and
projection are quite similar, and thus likely matched more cleanly. If
no viable image is found according to the same criteria as before, we
search the entire set of candidate images, selecting based on the same
scoring function from Figure \ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(c), where seams are now greatly reduced as
compared to Figure \ref{fig:compareAll}(b).

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Images for vertical planes are tilted, but their camera
    axes are more or less normal to their respective planes. (b)
    Camera axes for ceiling images are at large angles with respect to
    plane normals. (c) Wall images are cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, our data comes from a mobile backpack
system. Human operators can not carry the backpack in a perfectly
upright position and are bent forwards at 15 to 20 degrees with
respect to the vertical direction. As a result, cameras facing
sideways are head on with respect to vertical walls, while cameras
oriented towards the top or bottom of the backpack are at an angle
with respect to horizontal floors and ceilings. This is depicted in
Figures \ref{fig:projectionAngles}(a) and
\ref{fig:projectionAngles}(b). These oblique camera angles for
horizontal surfaces translate into textures that span large areas once
projected, as shown in Figure \ref{fig:projectionAngles}(b). Using the
tile-based texture mapping criteria from Figure
\ref{fig:scoringFunction}, such projections have highly varying scores
depending on the location of a tile on the plane. Thus, the tiling
approach in this section is a good choice for texturing floors and
ceilings, as it uses the parts of image projections that maximize
resolution and accuracy for their respective plane locations,
e.g. areas near point A and not near point B, in Figure \ref{fig:projectionAngles}(b).
